---
title: "lab-8-seandenny"
author: "Sean Denny"
date: "11/26/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Lab 8 - Linear Regression in R

```{r, load packages and data, message = FALSE}
library(tidyverse)
salmon <- read_csv("salmon_imports.csv")
```

```{r, exploratory scatterplot}

salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point()

salmon_scatter

#Data look to be more or less linear.
```

```{r, linear regression}

#model_name <- lm(y_variable ~ x_variable, data = df_name)

salmon_lr <- lm(salmon_fresh ~ year, data = salmon)

salmon_lr
```

Intercept is -14982940. Year is 7550.

Model equation:

Import(tons) = -14982940 + 7550.

• What does the slope mean in the context of this model?

ANSWER: 

• What does the y-intercept mean in the context of this model? Why is that concerning? What does this mean about extrapolating this model for past values?

ANSWER:

```{r, Model diagnostics}
par(mfrow = c(2,2))
plot(salmon_lr)
```

• Do residuals appear normally distributed?

ANSWER:

• Any concerns about heteroscedasticity or outliers?

ANSWER: 

```{r}
summary(salmon_lr)
```

• Does year significantly predict salmon imports?

ANSWER:

• What does the R2 value actually mean in words?

ANSWER:

• What proportion of the variance in salmon imports is NOT explained by year?

ANSWER:

```{r, final graph}
salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") +
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")

salmon_final_graph
```

```{r, make predictions using the model}
new_years <- data.frame(year = c(2022, 2024, 2026, 2028))
new_years
```

```{r}
future_predict <- predict(salmon_lr, newdata = new_years, interval = "confidence")
future_predict
```

```{r}
predictions <- data.frame(new_years, future_predict)

predictions
```

```{r, Find Pearsons r}
salmon_cor <- cor(salmon$salmon_fresh, salmon$year)

salmon_cor

#Value is 0.9735387
```

Pearson’s r for the year vs. salmon imports linear trend:

ANSWER:

Would you describe this as a weak/strong negative/positive correlation?

ANSWER: 

Using the doc *Communicating Results of Basic Linear Regression* (posted on GauchoSpace) as a guide, write a final 1 - 2 sentence statement describing the results of your linear regression and Pearson’s r findings.

ANSWER: 